---
title: "Overview of the Ecosystem for the Scotian Shelf 2021"
author: "Maritimes EBFM Working Group"
date: "5/28/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE,message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(data.table)
library(formattable)
library(sparkline)
library(ggplot2)
library(RColorBrewer)
library("magick")
library("webshot")
webshot::install_phantomjs()
```


# Overview of the Ecosystem for the Scotian Shelf 2021

## Introduction


### About this report
This report is for Fisheries and Oceans Canada fisheries managers and relevant stakeholders. The purpose of this report is to synthesize ecosystem information to inform an ecosystem approach, as highlighted in the modernized Fisheries Act (2019). It is based on the objectives of the [Maritimes Ecosystem Based Management Framework](http://publications.gc.ca/collections/collection_2020/mpo-dfo/Fs97-6-3368-eng.pdf) and ecosystem based objectives within IFMPs. The information in this report is organized into two sections: 1) to link relevant indicators to meet management objectives derived from the EBM framework, and 2) to provide sources of information relevant to the socio-ecological system (ie. the interconnected natural and social systems). 

### Report structure
The first section of this report reviews indicator trends and the status of the most recent year relative to a threshold (if available) or a long-term average (Table 1). We synthesize the results to outline potential implications to the region examined (connecting indicator status to management and why it is important). 

The second section of this report is a list of all indicators relevant to describing socio-ecological systems but that may not 1)have current and significant impacts to the region, 2)be a developed enough indicator for the region to be currently informative. This section contains links to data/reports/information sources. 

We emphasize that this is a guide to the most current information on socio-ecological systems and indicators and synthesis continue to develop. 


Table 1. Ecosystem-scale management objectives in the Scotian Shelf

```{r, echo=FALSE}

Table_01 <- tibble::tribble(
  ~"Management component", ~"Ecosystem Objective *", ~"Indicator",
    
      "Fishery",     "Maintain fishing pressure consistent with ecosystem's ability to recover
",     "Fishing pressure, Landings, Fleet structure, Effort by fleet
",
      "Non-target species",     "Fish sustainably while minimizing adverse impacts on non-target species
",     "Discards, Incidental catch, ghost gear
",
      "Other ocean users",     "",     "Aquaculture, Energy sector, Shipping and transport
",
      "Stakeholder engagement",     "",     "",
     
      "Habitat",     "Prevent unacceptable modification to habitat",     "Habitat types, habitat condition, connectivity
",
      "Oceanography",     "Key signals within the environment and ecosystem
",     "AMO, NAO, SST, Bottom Temperature, stratification, Climate anomalies
",
      "Productivity and trophic structure",     "Do not cause unacceptable reduction to productivity
",     "Primary and Secondary productivity, Size structure, PPR, Resource potential
",
      "Biodiversity",     "Do not cause unacceptable reduction in biodiversity
",     "Guild-level biomass, Species Diversity/Richness/Evenness, Charasmatic species, Invasive species
",
  "Indigenous Traditional Knowledge",     "",     "",
    
      "Commercial",     "Ensure ecosystem maintains economic prosperity and livelihoods
",     "Total revenue, Revenue components (Bennet indicator)
",
      "Non-commercial",     "Ensure ecosystem maintains economic prosperity and livelihoods
",     "Recreational fishing, bait
",
       
      "Healthy consumption",     "Ensure ecosystem continues to maintain happiness and wellbeing of individuals
",     "Contaminants, parasites
",
      "Benefits",     "Ensure ecosystem continues to maintain happiness and wellbeing of individuals
",     "Vulnerability index 
",
      "Equity",     "Ensure ecosystem continues to maintain happiness and wellbeing of individuals
",     "Women and underrepresented groups in industry
"
  )


kable(Table_01, digits = 3, row.names = FALSE, align = "l",
              caption = NULL) %>% 
  kable_paper("striped", full_width=F) %>%
  pack_rows("Ocean Users",1,4 ) %>%
  pack_rows("Ecological", 5,9) %>%
  pack_rows("Economic", 10,11) %>%
  pack_rows("Social and Cultural", 12, 14)%>%
  as_image(width=8)
```

## Section 1: Indicator performance for the Scotian Shelf

### Fishery Indicators

A summary of fisheries indicators (Fishing pressure, landings, fleet structure and effort by fleet) and how they have changed over time. What this might mean for the region and the last year. What do the indicators tell us? How could this affect management? What management decisions could be made?

Table 2: one way we could "preview" indicators NOT REAL DATA

```{r, echo=FALSE}

df <- data.frame(stringsAsFactors=FALSE,
          V1 = c("indicator", "Fishing pressure", "Landings", "Effort"),
          V2 = c(2000L, 100L, 600L, 50L),
          V3 = c(2001L, 250L, 500L, 80L),
          V4 = c(2002L, 370L, 300L, 70L),
          V5 = c(2003L, 399L, 400L, 80L),
          V6 = c(2004L, 580L, 500L, 100L),
          V7 = c(2005L, 600L, 120L, 150L)
)

df.names <- df[1,]
names(df) <- df.names
df <- df[-1,]

graph <- df %>% 
  group_by(indicator) %>% 
  gather(key=year, value=value, -indicator) %>% 
  summarise(graph=spk_chr(
    value, 
    chartRangeMin = 0,
    type="line"))

df2 <- left_join(df, graph, by=c("indicator"))


df2 %>%
  formattable::formattable(align=c("l")) %>% 
  as.htmlwidget() %>% 
  spk_add_deps()



```

```{r, echo=FALSE}
df = data.frame("Type" = c("bar", "line", "bullet", "pie", "tristate", "discrete"),
               Sparkline = c(as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "bar"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "line"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "bullet"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "pie"))), 
                             as.character(htmltools::as.tags(sparkline(c(-1,0,1,1,1,-1,0,2), type = "tristate"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "discrete")))))
out = as.htmlwidget(formattable(df))
out$dependencies = c(out$dependencies, htmlwidgets:::widget_dependencies("sparkline", "sparkline"))
out
```

Text about the individual indicators how to interpret them with more detailed graphs. Relevance to the management/ecological objective. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#catch data from NAFO 2J3KLNO
catchdataF<-read.csv(file="catchdataF.csv")
catchdataF %>%
  pivot_longer(-YEAR, names_to="species", values_to="catch")->catchdataF

 foragef<-filter(catchdataF, species==c("Capelin", "Atlantic.Herring", "Atlantic.Mackerel"))

groundf<-filter(catchdataF, species==c("Atlantic.Cod", "Yellowtail.Flounder", "Atlantic.Halibut"))

ffplot<-ggplot(data=foragef, aes(x=YEAR, y=catch, group=species))+
               geom_line(aes(color=species))+
               geom_point(aes(color=species))+

  scale_color_brewer() + theme_classic()
              
print(ffplot)

```

Figure 1. Landings of Forage fish in the Scotian Shelf (NOT REAL DATA)


## Section 2: List of available indicators for the Scotian Shelf region

```{r, echo=FALSE}
allindicators<-read.csv(file= 'Section2.csv')


kable(allindicators, digits = 3, row.names = FALSE, align = "l",
              caption = NULL) %>% 
  kable_paper("striped", full_width=F)%>%
 as_image(width=8)

```




## Glossary

Table x: Terminology and concepts

```{r, echo=FALSE}

glossary <- tibble::tribble(
  ~"Term", ~"Definition"  ,      
  "Indicator", "Quantitative biological, chemical, physical, social, or economic measurements that serve as proxies of the conditions of attributes of natural and socioeconomic systems.",
          "Ecosystem Attributes", "Characteristics that define the structure and function of the ecosystem.")
         
      


kable(glossary, digits = 2, row.names = FALSE, align = "l",
              caption = NULL) %>%
   kable_paper("striped", full_width=F) %>%
  as_image(width=8)
```




## Land Acknowledgement
We acknowledges we are in Mi’kma’ki – the unceded territory and ancestral homeland of the Mi’kmaq Nation.
