---
title: "Overview of the Ecosystem for the Scotian Shelf 2021"
author: "Maritimes EBFM Working Group"
date: "5/28/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default

---
```{r, echo=FALSE,message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(data.table)
library(formattable)
library(sparkline)
library("magick")
library("webshot")
webshot::install_phantomjs()
```


# Overview of the Ecosystem for the Scotian Shelf 2021

## Introduction


## About this report
This report is for Fisheries and Oceans Canada fisheries managers and relevant stakeholders. The purpose of this report is to synthesize ecosystem information to better meet fishery management objectives derived from the [Maritimes Ecosystem Based Management Framework](http://publications.gc.ca/collections/collection_2020/mpo-dfo/Fs97-6-3368-eng.pdf). The information in this report is organized into two sections: 1) to link relevant indicators to meet management objectives derived from the EBM framework, and 2) to provide sources of information relevant to the socio-ecological system. 

## Report structure
The first section of this report reviews indicator trends and the status of the most recent year relative to a threshold (if available) or a long-term average (Table 1). We synthesize the results to outline potential implications to the region examined (connecting indicator status to management and why it is important). 

The second section of this report is a list of all indicators relevant to describing socio-ecological systems but that may not 1)have significant impacts to the region, 2)be a developed enough indicator for the region to be currently informative. This section contains links to data/reports/information sources. 

We emphasize that this is a guide to the most current information on socio-ecological systems and indicators and synthesis continue to develop. 


Table 1. Ecosystem-scale management objectives in the Scotian Shelf

```{r, echo=FALSE}

Table_01 <- tibble::tribble(
  ~"Management component", ~"Ecosystem Objective *", ~"Indicator",
    
      "Fishery",     "Maintain fishing pressure consistent with ecosystem's ability to recover
",     "Fishing pressure, Landings, Fleet structure, Effort by fleet
",
      "Non-target species",     "Fish sustainably while minimizing adverse impacts on non-target species
",     "Discards, Incidental catch, ghost gear
",
      "Other ocean users",     "",     "Aquaculture, Energy sector, Shipping and transport
",
      "Stakeholder engagement",     "",     "",
     
      "Habitat",     "Prevent unacceptable modification to habitat",     "Habitat types, habitat condition, connectivity
",
      "Oceanography",     "Key signals within the environment and ecosystem
",     "AMO, NAO, SST, Bottom Temperature, stratification, Climate anomalies
",
      "Productivity and trophic structure",     "Do not cause unacceptable reduction to productivity
",     "Primary and Secondary productivity, Size structure, PPR, Resource potential
",
      "Biodiversity",     "Do not cause unacceptable reduction in biodiversity
",     "Guild-level biomass, Species Diversity/Richness/Evenness, Charasmatic species, Invasive species
",
  "Indigenous Traditional Knowledge",     "",     "",
    
      "Commercial",     "Ensure ecosystem maintains economic prosperity and livelihoods
",     "Total revenue, Revenue components (Bennet indicator)
",
      "Non-commercial",     "Ensure ecosystem maintains economic prosperity and livelihoods
",     "Recreational fishing, bait
",
       
      "Healthy consumption",     "Ensure ecosystem continues to maintain happiness and wellbeing of individuals
",     "Contaminants, parasites
",
      "Benefits",     "Ensure ecosystem continues to maintain happiness and wellbeing of individuals
",     "Vulnerability index 
",
      "Equity",     "Ensure ecosystem continues to maintain happiness and wellbeing of individuals
",     "Women and underrepresented groups in industry
"
  )


kable(Table_01, digits = 3, row.names = FALSE, align = "l",
              caption = NULL) %>% 
  kable_paper("striped", full_width=F) %>%
  pack_rows("Ocean Users",1,4 ) %>%
  pack_rows("Ecological", 5,9) %>%
  pack_rows("Economic", 10,11) %>%
  pack_rows("Social and Cultural", 12, 14)%>%
  as_image(width=8)
```

## Section 1: Indicator performance for the Scotian Shelf

Table 2: one way we could "preview" indicators

```{r, echo=FALSE}

df <- data.frame(stringsAsFactors=FALSE,
          V1 = c("indicator", "A", "B", "C"),
          V2 = c(2000L, 100L, 600L, 50L),
          V3 = c(2001L, 200L, 500L, 60L),
          V4 = c(2002L, 300L, 400L, 70L),
          V5 = c(2003L, 400L, 300L, 80L),
          V6 = c(2004L, 500L, 200L, 90L),
          V7 = c(2005L, 600L, 100L, 100L)
)

df.names <- df[1,]
names(df) <- df.names
df <- df[-1,]

graph <- df %>% 
  group_by(indicator) %>% 
  gather(key=year, value=value, -indicator) %>% 
  summarise(graph=spk_chr(
    value, 
    chartRangeMin = 0,
    type="line"))

df2 <- left_join(df, graph, by=c("indicator"))


df2 %>%
  formattable::formattable(align=c("l")) %>% 
  as.htmlwidget() %>% 
  spk_add_deps()


```


Table 3: other widget options?

```{r, echo=FALSE}
df = data.frame("Type" = c("bar", "line", "bullet", "pie", "tristate", "discrete"),
               Sparkline = c(as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "bar"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "line"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "bullet"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "pie"))), 
                             as.character(htmltools::as.tags(sparkline(c(-1,0,1,1,1,-1,0,2), type = "tristate"))), 
                             as.character(htmltools::as.tags(sparkline(c(1,2,7,6,5), type = "discrete")))))
out = as.htmlwidget(formattable(df))
out$dependencies = c(out$dependencies, htmlwidgets:::widget_dependencies("sparkline", "sparkline"))
out
```

## Section 2: List of available indicators for the Scotian Shelf region


II.	Key Signals within the Environment and Ecosystem: Monitor features that directly impact the sustainability of marine resources.
a.	Oceanographic Overview
i.	Physical
1.	North Atlantic Oscillation
2.	Atlantic Multi-decadal Oscillation
3.	Annual Temperature Cycle
4.	Sea Surface Temperature
5.	Cold Intermediate Layer
6.	Sea Ice
7.	Bottom and Deep Water Temperatures
8.	Runoff and Stratification
9.	Labrador Current Transport Index
ii.	Chemical
1.	pH and CO2
2.	Biochemical conditions
3.	Nutrients
4.	Runoff
b.	Productivity: Do not cause unacceptable reduction in productivity so the components can play their role in the functioning of the ecosystems.
i.	Ecosystem Structure and Functioning
1.	Lower Trophic Levels
a.	Primary Productivity
i.	Phytoplankton
b.	Secondary Productivity
i.	Zooplankton (biomass and size)
2.	Ecosystem Structure
a.	Food Web Disruptions
i.	Warm Water Species
ii.	Aquatic Invasive Species
b.	Biomass of Trophic Guilds and Forage Components
c.	Size Structure (fish community)
d.	Spatial
i.	Quantity Occupancy Aggregation
ii.	Area-based Changes in Thermal Habitat
iii.	Patchiness
3.	Structural Change
4.	Ecosystem Functioning
ii.	Resource Potential
1.	Resource Potential of the Community
2.	Resource Potential of the Fished Groups
3.	Resource Potential of Indicator Species
4.	Fishing Strategies and System Productivity
a.	Spatial occupancy
b.	Production Potential 
c.	Conservation of Biodiversity: Do not cause unacceptable reduction in biodiversity in order to preserve the structure and natural resilience of the ecosystem.
i.	Biodiversity
1.	Keystone and Indicator Species
a.	Plankton
i.	Phytoplankton
ii.	Zooplankton
b.	Marine Vegetation
c.	Marine Invertebrates
d.	Marine Fish
e.	Marine Mammals
i.	Cetaceans
ii.	Seals
f.	Seabirds
g.	Sea Turtles
2.	Species Richness
3.	Species Diversity
4.	Species Evenness
5.	Range expansion/contraction
ii.	Ecosystem Stability and Resistance to Perturbations
1.	Ecosystem Resistance
2.	Stability of the Trophic Community
3.	Stability of Community Biomass
III.	Habitat: Prevent unacceptable modification to habitat in order to safeguard the biological, physical and chemical properties of the ecosystem that allow species to complete their life cycle.
i.	Habitat 
1.	Biogenic
a.	Corals and Sponges
b.	Dominant Species
c.	Eelgrass
d.	Kelp
e.	Mudflats 
f.	Sand dollars
g.	Near-shore
h.	Off-shore
i.	Benthic
j.	Pelagic
k.	Intertidal vegetation
2.	Non-biogenic
a.	Rocky Shores
3.	Area and Structure
a.	Area/Extent
b.	Habitat Types
c.	Fractal Dimension
d.	Connectivity
4.	Habitat Condition
a.	Abiotic Properties
b.	Biotic Properties
c.	Dynamic Structure
d.	Water Condition
e.	Benthic Condition
5.	Prey
IV.	Key Signals within the Fishery: Maintain fishing pressure at levels consistent with the ecosystem's ability to recover from it.
i.	Fishing Pressure 
1.	Fishing Pressure on the Community 
2.	Fishing Pressure on Fished Groups
3.	Distribution of Fishing Pressure
4.	Landings of the Community
5.	Landings of Fished Groups
6.	Landings of Indicator Species
7.	Number and size of vessels by fleet
8.	Effort by fleet
V.	Social and Economic Considerations: Ensure that the ecosystem continues maintain the prosperity, happiness and wellbeing of individuals and communities that derive a livelihood from it. 
i.	Commercial 
1.	Total Revenue
2.	Revenue Components (Bennet Indicator)
3.	Stakeholder Engagement
4.	Overlap with Offshore Activities (wind lease areas)
5.	Oil and Gas Activities (exploration, spills)
6.	Pollution (litter occurrences and density from commercial fleets)
ii.	Non-commercial Fisheries
1.	Recreational Fishing Harvest and Effort
a.	Profit-based (e.g., tuna outfitters) (Alex Hanke)
b.	Socio-cultural (e.g., squid jigs)
c.	Bait (e.g., mackerel)
iii.	Healthy Consumption
1.	Contaminants
2.	Parasites
VI.	Management: Fish sustainably while minimizing adverse impacts on non-target species.
i.	Precautionary Approach Application
1.	Total Catch Indices
2.	Multi-species and/or environmental interactions
3.	Production Potential (single species)
ii.	Minimize Harmful Impacts of Fishing
1.	Level of Protection of Ecosystem
2.	Level of Protection of Exploited Species
iii.	Assess Significance of Incidental Mortality of Fishing
1.	Discards across fisheries
2.	Incidental catch of depleted/protected species
iv.	Area-based management 


## Glossary

Table x: Termanology and concepts

```{r, echo=FALSE}

glossary <- tibble::tribble(
  ~"Term", ~"Definition"  ,      
  "Indicator", "Quantitative biological, chemical, physical, social, or economic measurements that serve as proxies of the conditions of attributes of natural and socioeconomic systems.",
          "Ecosystem Attributes", "Characteristics that define the structure and function of the ecosystem.")
         
      


kable(glossary, digits = 2, row.names = FALSE, align = "l",
              caption = NULL) %>%
   kable_paper("striped", full_width=F) %>%
  as_image(width=8)
```




## Land Acknowledgement
We acknowledges we are in Mi’kma’ki – the unceded territory and ancestral homeland of the Mi’kmaq Nation.
